# Analytics of Managerial Economics Project 1 
Tan Tian Tze A0161067J

Executive summary (not more than half a page)
The aim of this study is to gain a better understanding of roasted coffee market in
Netherlands to increase the market share and profit of the ABC Coffee Roasters.
We analysed the coffee consumption data between 1990 to 1996 from the Dutch
Market to estimate the demand and supply functions. We estimated the elasticity
of demand to be -0.28661 which yields a monopolistic price of 4.48 times
Marginal cost. However, the market clearing price is between Marginal Cost and
Monopolistic price which indicates an oligopolistic Market. In addition, we see an
evidence of interdependent behaviour in the seasonal fluctuation in the supply.
Based on this analysis we believe a strategy based on selling differentiated
products and building brand loyalty would work best in this market.
Abstract
Quick highlights - 
Demand
demand side is fine. Just need to talk about tprice and income
Supply
supply side is failing, wprice is wierdly related to cprice (can be due to policy for wages), basic model assumptions might be wrong


Assumptions
Assumption of the demand function - tends to perfect competition, simple constant elasticity model, enough time for supplier and demand to react to price (level of data)

Demand function (not more than 1.5 pages)
Model comparison
Simple OLS vs TSLS
Defend choice of control variables and IV's
2 assumptions checks
Income vs Tprice
Why ignore oprice?
Why not use both tprice and income?
Hausman test for instruments

Supply function ( ideally 2 pages)
Model comparison
Defend choice of control variables and IV's
Problem about controls - Cprice not being significant in some cases, and -ve sign in other cases
Why wprice and bprice not working together? - try wprice, try bprice, try both
This could be because our original assumption about constant supply elasticity unsuitable
Another reason why not working could be because of oligopolistic coffee market in Netherlands, so supply doenst depend on price
Hausman proves choice of instruments to be fine - tried income, quarter(not significant) and tprice

Appendix

Log


## Estimating Coffee Demand and Supply Functions
maand - year and month of observation
year - year of observation
month - month of observation
qu - per capita consumption of roasted coffee in kg
cprice -  price of roasted coffee per kg in current guilders
tprice - price of per kg tea in current guilders
oprice - price index for other goods
income - income per capita in current guilders
q1 - season dummy 1
q2 - season dummy 2
q3 - season dummy 3
q4 - season dummy 4
bprice - price of coffee beans per kg in current guilders
wprice - price of labor per man hours (work 160 hours per month)

The price elasticity of demand and price elasticity of supply for roasted coffee.  Findings will contribute to firms’ understanding of how revenue would change due to price changes (i.e. price elasticity of demand), and how quickly and effectively firms could respond to market conditions, especially changes in price (i.e. price elasticity of supply) (Agarwal, 2016).

Agarwal, P. (2016). Price elasticity of demand. Intelligent Economist. Retrieved from: https://www.intelligenteconomist.com/price-elasticity-of-demand/

```{r}
rawdata <- read.csv("Project1Data.csv")
```
## Factors of production 
Calculating Input_price (factors of production) and % of labour and bean price of Input_price

```{r}
rawdata$input_price <- rawdata$wprice + rawdata$bprice 
rawdata$wprice_percent <- rawdata$wprice / rawdata$input_price
rawdata$bprice_percent <- rawdata$bprice / rawdata$input_price
```
## Adjustment for Inflation 
Controlling for inflation by dividing the price variable by oprice
cprice -  price of roasted coffee per kg in current guilders
tprice - price of per kg tea in current guilders
oprice - price index for other goods
income - income per capita in current guilders
bprice - price of coffee beans per kg in current guilders
wprice - price of labor per man hours (work 160 hours per month)
```{r}
#matrix of factors adjusted by inflation 
adj.data <- rawdata[ ,c(5,6,8,13,14)] / rawdata$oprice
# changing col names 
colnames(adj.data) <- paste(colnames(adj.data),"adj", sep = "_")

combine.adj <- cbind(rawdata, adj.data)
```
## Log transform variables
Since we are assuming an exponential demand function with constant elasticity, we need to log transform all the necessary variables

```{r}
log.transform <- log((combine.adj[,-c(1:3, 9:12)]))
colnames(log.transform) <- paste("ln", colnames(log.transform), sep = "_")
combine.adj.ln <- cbind(combine.adj, log.transform)
```
 
## Demand function estimation
Now running a *simple OLS* between cprice and qu, to get the control variables 
```{r}
lm1.dd <- lm(ln_qu ~ ln_cprice + ln_tprice + ln_incom, data = combine.adj.ln)
summary(lm1.dd)
```
Estimate Std. Error t value Pr(>|t|)   
(Intercept)  -4.6814     2.1808  -2.147   0.0348 * 
ln_cprice    -0.3581     0.1321  -2.711   0.0082 **
Multiple R-squared:  0.1162,	Adjusted R-squared:  0.08304 

price is an endogenous variable, and not exogenous, we need to run a TSLS, to get the accurate prediction of demand elasticity
  
## Demand Two stage Least Square        
The first step is finding an estimate of Cprice using intrument variables and controlling for demand using control variables

```{r}
lm2.dd <- lm(ln_cprice_adj ~ ln_bprice_adj + ln_incom_adj + ln_tprice_adj, data = combine.adj.ln)
summary(lm2.dd)
# ln_bprice_adj  0.48654    0.01786  27.243   <2e-16 ***
# Multiple R-squared:  0.9197,	Adjusted R-squared:  0.9167          

# Use fitted values of price from lm2.dd and save in a column, run a 2nd step of regression, to correctly get estimate
combine.adj.ln$cprice_pred2_dd <- lm2.dd$fitted.values

# Run the 2nd stage OLS with the predicted values of coffee 
lm2.dd.pred <- lm(ln_qu ~ cprice_pred2_dd + ln_incom_adj + ln_tprice_adj, data = combine.adj.ln)
summary(lm2.dd.pred)
```
(Intercept)     -8.85374    3.71823  -2.381  0.01963 * 
cprice_pred2_dd -0.29798    0.09884  -3.015  0.00345 **
ln_incom_adj     1.01246    0.34886   2.902  0.00479 **

Multiple R-squared:  0.1451,	Adjusted R-squared:  0.1131 

## Check correlation between Instrument variables with residuals in model lm2.dd.pred

```{r}
hausman.test.dd <- cbind(as.data.frame(combine.adj.ln$ln_bprice_adj), as.data.frame(lm2.dd.pred$residuals))

colnames(hausman.test.dd) <- c("ln_bprice","residuals")

hausman_test <- lm(residuals ~ ln_bprice, data = hausman.test.dd)
summary(hausman_test)
```
No correlation seen. Independence can be assumed.
Multiple R-squared:  2.312e-29,	Adjusted R-squared:  -0.0122 

## Supply function estimation
*Simple OLS of Supply* between cprice and qu, to get the control variables
qu - per capita consumption of roasted coffee in kg
cprice -  price of roasted coffee per kg in current guilders
bprice - price of coffee beans per kg in current guilders
wprice - price of labor per man hours (work 160 hours per month)
```{r}
lm1.ss <- lm(ln_qu ~ ln_cprice + ln_bprice + ln_wprice + q1 + q2 + q3, data = combine.adj.ln)
summary(lm1.ss)
```
Multiple R-squared:  0.2736,	Adjusted R-squared:  0.217 
Not great, seemingly no relationship. 

## TSLS on Supply function for Coffee
Firstly, predict (fitted values) Cprice using instrument variables and controlling for demand using control variables
```{r}
lm2.ss <- lm(ln_cprice_adj ~ ln_incom_adj + ln_bprice_adj, data = combine.adj.ln)
summary(lm2.ss)
```

### Predicted cprice using lm2.ss model

```{r}
combine.adj.ln$cprice_pred_ss <- lm2.ss$fitted.values
```

## 2nd Stage OLS on Supply using cprice_pred_ss

```{r}
lm2.ss.pred <- lm(ln_qu ~ cprice_pred_ss + ln_bprice_adj + q1 + q2 + q3, data = combine.adj.ln)
summary(lm2.ss.pred)
```
ln_bprice_adj  -0.69149    0.37837  -1.828 0.071442 .  
q1             -0.11158    0.03179  -3.510 0.000747 ***
q2             -0.08833    0.03017  -2.928 0.004468 ** 
q3             -0.10846    0.03073  -3.529 0.000702 ***
Multiple R-squared:  0.2827,	Adjusted R-squared:  0.2367 

### Correlation of Instrument variables chosen with residuals

```{r}
hausman.test.ss <- cbind(as.data.frame(combine.adj.ln$ln_cprice_adj), as.data.frame(lm2.ss.pred$residuals))

colnames(hausman.test.ss) <- c("ln_cprice","residuals")

hausman_test_supply <- lm(residuals ~ ln_cprice, data = hausman.test.ss)
summary(hausman_test_supply)
```
No correlation seen. Independence can be assumed.
Multiple R-squared:  3.881e-05,	Adjusted R-squared:  -0.01216 

```{r}

```


```{r}
sessionInfo()
```

